Supplementary Information

### 2021-02-16 writing/transect_paper.Rmd
more revisions

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@nilsson2012].
<!--R4: Near L114 - More refs to consider: See Ashelford et al., 2005 Appl Env Microbiol 71: 7724,
Bidartondo et al., 2008 Science 319:1616, but also LeRay et al., 2019 PNAS 116: 22651 for a
more recent take-->
Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent.
Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
<!--R4: see Nilsson et al., 2005 BMC Bioinformtics 6: 178-->
The Unite database is a more curated <!--R5: More curated?-->attempt to include all publically available ITS sequences (800\ 000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
The Ribosomal Data Project [RDP, @cole2014] hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
the Warcup ITS training set, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [RDP-LSU, @liu2012].
Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.

Assigning taxonomy to unknown sequences using BLAST requires *a priori* choice of similarity thresholds for different taxonomic ranks.
Several algorithms specifically designed for taxonomic assignment have been published which instead use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017;@palmer2018].
However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic [@nilsson2016top;@Tedersoo2018].
<!--R4: consider mentioning Nilsson et al., 2019 Nuc Acids Res 47: D259 on the naming of dark
taxa in UNITE see: taxon hypotheses (THs)-->

@@

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
<!--R5: Add citation for “turnover scale”-->
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups,
and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
or by fitting a function to an empirical distance-decay curve of community dissimilarity vs. distance [@legendre2012].

@@

Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Fabaceae tr. Amherstieae).
At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag. <!--R4: how was the knife sterilized, how were the samples homogenized? Preserved at what
temperature? Please address in main text.-->
A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).

## DNA amplification and sequencing
DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and an equimolar mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
See Supplementary Methods \@ref(DNA-supplement) for PCR and library preparation protocols.
<!--R4: I always recommend that full methods be in the main text, not in the supplement, it only
takes a few sentences to describe PCR protocols at least.-->

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods \@ref(long-numbers)).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.

@@

<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.
Regular indentations at the start of each paragraph would make this more clear.-->
Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.

<!--R4: this looks like maybe a new paragraph? It should be connected to last paragraph.-->
To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].

@@

Taxonomic annotations of the RDP-LSU training set version 11.5 [@liu2012;@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods \@ref(reannotate)).
Primary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three popular algorithms:
the RDPC [@wang2007] as implemented in DADA2;
SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
and IDTAXA [@murali2018].
A relatively lax confidence threshhold of 50% was used for all three algorithms.
<!--R4: a relatively lax bootstrap support cutoff of 50% would indicate classification accuracy of
what? 75% accuracy, more, less, does it vary across ITS and LSU? I suggest this be addressed in
the main text, in https://rdp.cme.msu.edu/download/posters/fungalITSreport_062014.pdf see
Fig 3. -->
<!--R5: This seems very low. I think that some explanation for such a low value is needed to give a rationale. This is basically a coin toss so that is not exactly giving great confidence in the results-->
Each full-length ASV was thus given up to nine primary taxonomic assignments (three references $\times$ three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
The PHYLOTAX algorithm resolves conflicts and makes additional assignments based on an input tree (here the ML tree generated above) and taxonomic identifications from one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.

ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the preliminary assignments [@huson2007].
This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [e.g., MEGAN and LCAClassifier, @huson2016;@lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as "Consensus".

@@

To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
Each of these distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.

Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary Methods \@ref(spatiotemporal-supplement)).

@@

Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads (Figure \@ref(fig:sample-depth)).
Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, and taxonomic assignment.
<!--R5: So make a statement about how many samples are left in the analyses after these were excluded? Seems like it may be a very low number of samples here-->

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
Sequencing with PacBio yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent and Illumina yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads ($\le 1$ expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
Demultiplexing, primer trimming, and quality filtering reduced the read totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
<!--R5: But how does that impact quality? Seems like this would reduce quality significantly -->
Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of denoised ITS2 reads from the long-amplicon PacBio dataset.
Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed `r readd(bioinf_table)[c("long", "ITS2"), c("Ion Torrent_Short_reads", "Illumina_Short_reads", "PacBio_Short_reads")] %>% {glue_collapse(percent(.["long",]/.["ITS2",], format = "d", digits = 0), ", ", last = ", and ")}` of denoised reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \@ref(tab:bioinfo)).

@@

(ref:venn) Shared richness and abundance of ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
In each region, the ASV/OTU richness is given above, while the relative abundance of reads represented by these ASVs/OTUs in each sequencing strategy are shown below in the order PacBio/Illumina/Ion Torrent (*a*, *b*), or long/short (*c*, *d*).
For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.
<!--R4: it would be helpful if you drew a line down the middle of this figure then labelled the
left side ASVs and the right side OTUs. Instead of ‘abundance’ suggest you change to ‘read
number or read abundance’. Is this figure based on rarefied, normalized data? Indicate in the
Text.-->
<!--R5: again, were these analyses conducted on the entire (i.e. fungal and non-fungal) dataset, or just the fungal community?-->

```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", fig.height = 4, fig.align="center", results="asis"}
ggpubr::ggarrange(readd(vennplot_tech_ASV), readd(vennplot_tech_OTU), ncol = 2)
grid::grid.legend(
  labels =  c("PacBio Short", "Illumina", "Ion Torrent"),
  nrow = 1,
  pch = 21,
  gp = grid::gpar(fill = c("grey85", "lightblue", "lightcoral")),
  vp = grid::viewport(y = unit(0.05, "npc"))
)

@@

ggpubr::ggarrange(readd(vennplot_amplicon_ASV), readd(vennplot_amplicon_OTU), ncol = 2)
grid::grid.legend(
  labels =  c("Long amplicon", "Short amplicon"),
  nrow = 1,
  pch = 21,
  gp = grid::gpar(fill = c("white", "slateblue1")),
  vp = grid::viewport(y = unit(0.15, "npc"))
)
```

>>>>

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Although the public INSDC databases [@karsch-mizrachi2018] are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations [@bidartondo2008;@nilsson2006;@steinegger2020] as well as sequences of poor technical quality [@ashelford2006;@nilsson2012].
Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent [@nilsson2005]. 
Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
The Unite database is an attempt to include all publically available high-quality ITS sequences (800\ 000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes [@nilsson2019a], where efforts have been made to correct incorrect annotations and exclude low-quality sequences [@abarenkov2018a].
The Ribosomal Data Project [RDP, @cole2014] hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
the Warcup ITS training set, containing 18\ 000 manually curated fungal ITS sequences [@deshpande2016], and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups [RDP-LSU, @liu2012].
Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.

Assigning taxonomy to unknown sequences using BLAST requires *a priori* choice of similarity thresholds for different taxonomic ranks.
Several algorithms specifically designed for taxonomic assignment have been published which instead use information about variability within different taxa in the reference database to assign unknown sequences, along with confidence estimates for these assignments, including the RDP Classifier [RDPC, @wang2007], SINTAX [@edgar2016a] and IDTAXA [@murali2018a] among others.
In addition, methods have been published which integrate predictions from multiple algorithms to increase the reliability of assignments [@somervuo2016;@gdanetz2017;@palmer2018].
However, all sequence-similarity based approaches are dependent on high taxonomic coverage in the reference database, making the placement of novel or undersampled groups problematic [@nilsson2016top;@Tedersoo2018].

@@

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool [@legendre2012].
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups,
and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
or by fitting a function to an empirical distance-decay curve of community dissimilarity vs. distance [@legendre2012].

@@

Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30\ km apart in the _Forêt Classée de l'Ouémé Supérieur_ (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands [@olson2001;@yorou2014] dominated by the ECM host tree _Isoberlinia doka_ (Fabaceae tr. Amherstieae).
At each site, 25\ soil samples were collected along a single 24\ m linear transect at intervals of 1\ m in May 2015.
One third of the sample locations (3\ m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5\ cm\ × 5\ cm\ × 5\ cm was extracted with an ethanol sterilized knife blade and homogenized by manually breaking up clumps and mixing inside a plastic zipper bag.
A subsample of approximately 250\ mg was preserved and returned to the laboratory for extraction (see Supplementary Methods \@ref(DNA-supplement) for preservation and extraction methods).

## DNA amplification and sequencing
DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure\ \@ref(fig:regions)a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
The short amplicon (approximately 300\ bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 [@ihrmark2012] as the forward primer and an equimolar mix of ITS4 [@white1990amplification] and ITS4a [@urbina2016] as the reverse primer.
The long amplicon (approximately 1500\ bp) targeted the full ITS region including the 5.8S rDNA and approximately 950\ bp at the 5' end of the LSU, including the first three variable regions (Figure\ \@ref(fig:regions)a), using ITS1 [@white1990amplification] as the forward primer and LR5 [@vilgalys1990] as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.
See Supplementary Methods \@ref(DNA-supplement) for PCR and library preparation protocols.

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods \@ref(long-numbers)).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU [@michot1984], and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure\ \@ref(fig:regions)a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, implemented in the new R packages `LSUx` (splitting reads into homologous regions; https://github.com/brendanf/LSUx) and `tzara` (reassembling regions and generating full-length consensus ASVs; https://github.com/brendanf/tzara) and detailed in Supplementary Methods \@ref(denoising-supplement), for all of the PacBio and Ion Torrent datasets.
Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.
To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97% similarity using VSEARCH v2.9.1 [@rognes2016].

@@

@@

Taxonomic annotations of the RDP-LSU training set version 11.5 [@liu2012;@cole2014] and Warcup ITS training set [@deshpande2016] were mapped to a uniform taxonomic classification system (see Supplementary Methods \@ref(reannotate)).
Primary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three popular algorithms:
the RDPC [@wang2007] as implemented in DADA2;
SINTAX [@edgar2016a] as implemented in VSEARCH v2.9.1 [@rognes2016];
and IDTAXA [@murali2018].
A relatively lax confidence threshold of 50% was used for all three algorithms, in order to increase the amount of input for  consensus algorithms.
Each full-length ASV was thus given up to nine primary taxonomic assignments (three references $\times$ three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

For full-length long-amplicon ASVs, the primary taxonomic assignments were refined with the new algorithm PHYLOTAX.
The PHYLOTAX algorithm resolves conflictsm one or more primary assignment methods (see Figure\ \@ref(fig:phylotax) and Supplementary Methods \@ref(phylotax)).
It is available in the new R package `phylotax` at https://github.com/brendanf/phylotax.

ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different primary assignments at each rank, resulting in a consensus assignment equivalent to the "last common ancestor" of the primary assignments [@huson2007].
This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits [e.g., MEGAN and LCAClassifier, @huson2016;@lanzen2012] or k-mer similarity scores [mothur's k-nearest neighbor method, @schloss2009], but here is used to resolve conflicts between assignments from different algorithms and databases.
Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as "Consensus".

@@

To estimate turnover scale, ecological community dissimilarity matrices were calculated using the ASV/OTU based Bray-Curtis metric [@bray1957, for both long and short amplicons] and the phylogenetically based weighted UniFrac metric [@lozupone2005;@lozupone2007, for only long amplicons] in `phyloseq` version `r packageVersion("phyloseq")`.
Dissimilarities were based on relative read abundance within each sample.
Samples were not rarefied to a standard sequencing depth within datasets, as both the Bray-Curtis dissimilarity and the UniFrac metric are robust to unequal sampling depths [@mcmurdie2014].
In addition, we did not standardize sequencing depth between datasets, because this would remove one of the potential benefits of Illumina and Ion Torrent relative to PacBio.

Each of the distance matrices was used to calculate a Mantel correlogram with a 1\ m bin size for distances in the range of 0--12\ m, i.e., half the maximum separation present in the dataset.
Separate correlograms were drawn for samples taken during the same year and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.
Additionally, empirical spatiotemporal distance-decay curves were generated by plotting mean community dissimilarity as a function of spatial distance and time lag, and fit to an exponential model of the form given by @legendre2012 using the `nls` function in R (Supplementary Methods \@ref(spatiotemporal-supplement)).
Spatiotemporal analyses were performed on the full recovered fungal community after removal of non-fungal sequences, and on the ECM community.

@@

Samples from Ang in 2015 yielded low quantities of DNA, poor PCR performance, and ultimately very few sequencing reads, especially in the long amplicon library, where only one sample produced more than 100\ reads (Figure \@ref(fig:sample-depth)).
Consequently, Ang samples were excluded from spatial analysis, although they were retained for denoising, phylogenetic reconstruction, taxonomic assignment, and all non-spatial analyses.
Spatial analyses were based on the remaining 34 samples for Illumina and 30 samples each for the PacBio Long and Short amplicons.

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
Sequencing with PacBio yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent and Illumina yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads ($\le 1$ expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
Demultiplexing, primer trimming, and quality filtering reduced the read totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
Extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) reduced the loss of long amplicon PacBio reads to only `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")`, comparable to Illumina.
Application of `tzara` resulted in `r readd(bioinf_table)["long", "PacBio_Long_ASVs"]` reconstructed long-amplicon ASVs, representing `r percent(readd(bioinf_table)["long", "PacBio_Long_reads"] / readd(bioinf_table)["ITS2", "PacBio_Long_reads"], format = "f", digits = 0)` of denoised ITS2 reads from the long-amplicon PacBio dataset.
Mapping identical ITS2 ASVs from the short and long amplicon datasets allowed `r readd(bioinf_table)[c("long", "ITS2"), c("Ion Torrent_Short_reads", "Illumina_Short_reads", "PacBio_Short_reads")] %>% {glue_collapse(percent(.["long",]/.["ITS2",], format = "d", digits = 0), ", ", last = ", and ")}` of denoised reads from the Ion Torrent, Illumina, and PacBio short amplicon datasets, respectively, to be assigned to a long amplicon ASV (Table \@ref(tab:bioinfo)).

@@

(ref:venn) Shared richness and abundance of ITS2-based ASVs (*a*, *c*) and 97% OTUs (*b*, *d*) between different sequencing technologies from the same short amplicon library (*a*, *b*), and between long and short amplicon libraries (*c*, *d*).
In each region, the ASV/OTU richness is given above, while the relative abundance of reads represented by these ASVs/OTUs in each sequencing strategy are shown below in the order PacBio/Illumina/Ion Torrent (*a*, *b*), or long/short (*c*, *d*).
For short amplicons in *c* and *d*, ASV/OTU counts reflect detection by any of the three technologies, and read counts represent the mean fraction of reads across the three technologies.  Analyses performed on pooled ASVs/OTUs from all samples.

```{r venn, fig.cap="(ref:venn)", fig.scap = "(ref:venn-s)", fig.show="hold", fig.height = 4, fig.align="center", results="asis"}
line = ggplot() +
  geom_vline(xintercept = 0) +
  theme_pubclean() +
  theme(axis.title = element_text(color = "white"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank()
        )
ggpubr::ggarrange(
  readd(vennplot_tech_ASV),
  line,
  readd(vennplot_tech_OTU),
  ncol = 3,
  widths = c(1,0.02,1),
  labels = c("ASVs", "", "OTUs"),
  label.x = 0.5,
  hjust = 0.5,
  font.label = list(size = 16)
)
grid::grid.legend(
  labels =  c("PacBio Short", "Illumina", "Ion Torrent"),
  nrow = 1,
  pch = 21,
  gp = grid::gpar(fill = c("grey85", "lightblue", "lightcoral")),
  vp = grid::viewport(y = unit(0.05, "npc"))
)

@@

ggpubr::ggarrange(
  readd(vennplot_amplicon_ASV),
  line,
  readd(vennplot_amplicon_OTU),
  ncol = 3,
  widths = c(1,0.02,1)
)
grid::grid.legend(
  labels =  c("Long amplicon", "Short amplicon"),
  nrow = 1,
  pch = 21,
  gp = grid::gpar(fill = c("white", "slateblue1")),
  vp = grid::viewport(y = unit(0.15, "npc"))
)
```

We compared the unique ASVs and OTUs shared between datasets from different sequencing strategies, and the number of reads represented by these ASVs and OTUs in each strategy.
The majority of abundant ASVs and OTUs were detected by all sequencing strategies used (Tables\ \@ref(tab:venn-table-asv) and \@ref(tab:venn-table-otu)).
The  short amplicon ASVs shared between all sequencing technologies represented `r text_list(percent(colSums(mutate_all(readd(venn_ASV)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` <!-- make this match--> of the reads for PacBio, Illumina, and Ion Torrent, respectively (Figure\ \@ref(fig:venn)a).
When differences at the intra-species scale were removed by clustering the ASVs into 97% OTUs, the number of OTUs shared between all three technologies increased to `r sum(readd(venn_OTU)[c("0111", "1111"),"OTUs"])`, representing `r text_list(percent(colSums(mutate_all(readd(venn_OTU)[c("0111", "1111"), c("reads_frac_pb_483", "reads_frac_SH-2257", "reads_frac_is_057")], as.numeric), na.rm = TRUE)))` of reads, respectively (Figure\ \@ref(fig:venn)b).
In particular, the majority of the `r sum(readd(venn_ASV)[c("1001","0001"),"ASVs"])` unique Ion Torrent ASVs were found to be shared with other sequencing technologies upon OTU clustering.
ASVs unique to the Ion Torrent dataset made up `r percent(sum(as.numeric(readd(venn_ASV)[c("1001","0001"),"reads_frac_is_057"])))` of reads in that dataset, but only `r percent(sum(as.numeric(readd(venn_OTU)[c("1001","0001"),"reads_frac_is_057"])))` belonged to a unique OTU after clustering.
In contrast, `r percent(readd(venn_ASV)["1000", "reads_frac_pb_500"])` of reads in the long PacBio dataset belonged to ASVs whose ITS2 region was unique to that dataset (Figure\ \@ref(fig:venn)c), and the fraction only reduced to `r percent(readd(venn_OTU)["1000", "reads_frac_pb_500"])` after clustering the ITS2 regions into OTUs (Figure\ \@ref(fig:venn)d).
The taxonomic identity of these ASVs is discussed below.

### 2021-02-16 writing/transect_supplement.Rmd
more revisions

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
readd(taxon_plot_ecm)
```

>>>>

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_plot_ecm"), more.cache = datasets}
readd(taxon_plot_ecm)
```
